---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vinereg
[![R-CMD-check](https://github.com/tnagler/vinereg/actions/workflows/R-CMD-check.yaml/badge.svg?branch=main)](https://github.com/tnagler/vinereg/actions/workflows/R-CMD-check.yaml)
[![Coverage status](https://codecov.io/gh/tnagler/vinereg/branch/main/graph/badge.svg)](https://app.codecov.io/github/tnagler/vinereg?branch=main)
[![CRAN status](https://www.r-pkg.org/badges/version/vinereg)](https://cran.r-project.org/package=vinereg)

Fit D-vine copula regression models for conditional mean and quantile
prediction with continuous or discrete variables.

## How to install

- the stable release from CRAN:

    ``` r
    install.packages("vinereg")
    ```

- the latest development version:

    ``` r
    # install.packages("remotes")
    remotes::install_github("tnagler/vinereg", build_vignettes = TRUE)
    ```

## Functionality

`vinereg` provides:

- automatic covariate selection or a user-specified D-vine order;
- parametric and nonparametric pair-copula families;
- conditional mean, quantile, distribution, and density predictions;
- support for continuous, ordered, and unordered variables; and
- standard model summaries, information criteria, and marginal effect plots.

See the [package website](https://tnagler.github.io/vinereg/) for the complete
reference and worked examples.

## Example

```{r, warning=FALSE}
set.seed(5)
library(vinereg)
data(mtcars)

# declare factors and discrete variables
for (var in c("cyl", "vs", "gear", "carb"))
    mtcars[[var]] <- as.ordered(mtcars[[var]])
mtcars[["am"]] <- as.factor(mtcars[["am"]])

# fit model
(fit <- vinereg(mpg ~ ., family_set = "nonpar", data = mtcars))

summary(fit)
AIC(fit)

# show marginal effects for all selected variables
plot_effects(fit)

# predict mean and median
head(predict(fit, mtcars, alpha = c(NA, 0.5)), 4)
```


## Vignettes

For more examples, have a look at the vignettes with

``` r
vignette("abalone-example", package = "vinereg")
vignette("bike-rental", package = "vinereg")
```

### References

Kraus and Czado (2017). D-vine copula based quantile regression.
*Computational Statistics \& Data Analysis*, 110, 1-18. 
[link](https://doi.org/10.1016/j.csda.2016.12.009), 
[preprint](https://arxiv.org/abs/1510.04161)

Schallhorn, N., Kraus, D., Nagler, T., Czado, C. (2017). D-vine quantile 
regression with discrete variables. arXiv preprint,
[preprint](https://arxiv.org/abs/1705.08310).
