---
title: "Example usage of the vinereg package"
author: "Daniel Kraus and Claudia Czado"
date: "September 2017"
output: 
    rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{Example usage of the vinereg package}
    %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
This file contains the source code of an exemplary application of the D-vine 
copula based quantile regression approach implemented in the R-package *vinereg*
and presented in Kraus and Czado (2017): 
*D-vine copula based quantile regression*, 
Computational Statistics and Data Analysis, 110, 1-18. 
Please, feel free to address questions to <daniel.kraus@tum.de>.
  
# Load required packages 
```{r, message = FALSE}
# install.packages("devtools")
# devtools::install_github("tnagler/vinereg")
library(vinereg) 
library(ggplot2)
library(dplyr)
library(tidyr)
```



# Data analysis


We consider the data set `abalone` from the UCI Machine Learning Repository (https://archive.ics.uci.edu/ml/datasets/abalone) and focus on the female
sub-population. In a first application we only consider continuous variables and
fit models to predict the quantiles of weight (`whole`) given the predictors 
`length`, `diameter`, and `height`. 

## Load and clean data

```{r,}
data(abalone, package = "PivotalR")
abalone_f <- abalone %>%
    dplyr::filter(sex == "F") %>%        # select female abalones
    dplyr::select(-id, -sex) %>%         # remove id and sex variables
    dplyr::filter(height < max(height))  # remove height outlier
```

```{r,, fig.width=7, fig.height=6}
pairs(abalone_f, pch = ".")
```


# D-vine regression models

## Parametric D-vine quantile regression

We consider the female subset and fit a parametric regression D-vine for the
response weight given the covariates len, diameter and height (ignoring the 
discreteness of some of the variables). The D-vine based model is selected 
sequentially by maximizing the conditional log-likelihood of the response 
given the covariates. Covariates are only added if they increase the (possibly 
AIC- or BIC-corrected) conditional log-likelihood.

We use the function `vinereg()` to fit a regression D-vine for predicting the 
response weight given the covariates `length`, `diameter`, and `height`. The 
argument `family_set` determines how the pair-copulas are estimated. We will
only use one-parameter families and the *t* copula here. The 
`selcrit` argument specifies the penalty type for the conditional 
log-likelihood criterion for variable selection.

```{r}
fit_vine_par <- vinereg(
    whole ~ length + diameter + height, 
    data = abalone_f, 
    selcrit = "aic"
)
```

The result has a field `order` that shows the selected covariates and their 
ranking order in the D-vine.
```{r,}
fit_vine_par$order
```

The field `vine` contains the fitted D-vine, where the first variable 
corresponds to the response. The object is of class `"vinecop_dist"` so we can
use `rvineocpulib`'s functionality to summarize the model
```{r,}
summary(fit_vine_par$vine)
```

We can also plot the contours of the fitted pair-copulas.
```{r,, fig.width=7, fig.height=7}
contour(fit_vine_par$vine)
```


## Estimation of corresponding conditional quantiles

In order to visualize the predicted influence of the covariates, we plot the 
estimated quantiles arising from the D-vine model at levels 0.1, 0.5 and 0.9 
against each of the covariates. 


```{r,}
# quantile levels
alpha_vec <- c(0.1, 0.5, 0.9) 
```

We call the `fitted()` function on `fit_vine_par` to extract the fitted values
for multiple quantile levels. This is equivalent to predicting the quantile at
the training data. The latter function is more useful for out-of-sample 
predictions.

```{r,}
pred_vine_par <- fitted(fit_vine_par, alpha = alpha_vec)
# equivalent to:
# predict(fit_vine_par, newdata = abalone.f, alpha = alpha_vec)
head(pred_vine_par)
```

To examine the effect of the individual variables, we will plot the predicted
quantiles against each of the variables. To visualize the relationship more 
clearly, we add a smoothed line for each quantile level. We will use 
this kind of plot more than once, so it's a good idea to write a small function 
for it. 

```{r,}
plot_marginal_effects <- function(covs, preds) {
    cbind(covs, preds) %>%
        tidyr::gather(alpha, prediction, -seq_len(NCOL(covs))) %>%
        dplyr::mutate(prediction = as.numeric(prediction)) %>%
        tidyr::gather(variable, value, -(alpha:prediction)) %>%
        ggplot(aes(value, prediction, color = alpha)) +
            geom_point(alpha = 0.15) + 
            geom_smooth(se = FALSE) + 
            facet_wrap(~ variable, scale = "free_x") +
            ylab(quote(q(y* "|" * x[1] * ",...," * x[p]))) +
            xlab(quote(x[k])) +
            theme(legend.position = "bottom")
}
```
```{r, fig.width=7, fig.height=4}
plot_marginal_effects(abalone_f[, 1:3], pred_vine_par)
```

The fitted quantile curves suggest a non-linear effect of all three variables.

## Comparison to the benchmark model: linear quantile regression

This can be compared to linear quantile regression:
```{r,, fig.width=7, fig.height=6}
pred_lqr <- pred_vine_par
for (a in seq_along(alpha_vec)) {
    my.rq <- quantreg::rq(
        whole ~ length + diameter + height, 
        tau = alpha_vec[a], 
        data = abalone_f
    )
    pred_lqr[, a] <- quantreg::predict.rq(my.rq)
}
plot_marginal_effects(abalone_f[, 1:3], pred_lqr)
```

## Nonparametric D-vine quantile regression

We also want to check whether these results change, when we estimate the 
pair-copulas nonparametrically.

```{r,, fig.width=4.6, fig.height=4.6}
fit_vine_np <- vinereg(
    whole ~ length + diameter + height,
    data = abalone_f,
    family_set = "nonpar",
    selcrit = "aic"
)
fit_vine_np$order
contour(fit_vine_np$vine)
```

Now only the length and height variables are selected as predictors. Let's have 
a look at the marginal effects.

```{r, fig.width=7, fig.height=4}
pred_np <- predict(fit_vine_np, newdata = abalone_f, alpha = alpha_vec)
plot_marginal_effects(abalone_f[, 1:3], pred_np)
```

The effects look similar to the parametric one, but slightly more wiggly. Note
that even the diameter was not selected as a covariate, it's marginal effect 
is captured by the model. It just does not provide additional information when
height and length are already accounted for.

## Discrete D-vine quantile regression

To deal with discrete variables, we use the theory which is developed in 
Nagler (2017) and applied to D-vine quantile regression in 
Schallhorn et al. (2017). For the estimation the discrete variable(s) are 
transformed to continuous ones by jittering. 

We let `vinereg()` know that a variable is discrete by declaring it `ordered`.

```{r,, fig.width=4.7, fig.height=4}
abalone_f$rings <- as.ordered(abalone_f$rings)
fit_disc <- vinereg(
    rings ~ .,
    data = abalone_f,
    family_set = "nonpar", 
    selcrit = "aic"
)
fit_disc$order
contour(fit_disc$vine)
pred_disc <- predict(fit_disc, newdata = abalone_f, alpha = alpha_vec)
plot_marginal_effects(abalone_f[, fit_disc$selected_vars], pred_disc)
```


# References
Kraus and Czado (2017), **D-vine copula based quantile regression**, *Computational Statistics and Data Analysis, 110, 1-18*

Nagler (2017), **A generic approach to nonparametric function estimation with mixed data**, *arXiv preprint*

Schallhorn, Kraus, Nagler and Czado (2017), **D-vine quantile regression with discrete variables**, *arXiv preprint*
